---
title: "Higher Ed Labor"
author: "Tom Ben-Shahar"
date: "February 2025"
output: pdf_document
---

Describe your issue very briefly here. Then show it with a minimal, self-contained example in the following R chunk.

1 - Professor
2 - Associate Professor
3 - Assistant Professor
4 - Instructor
5 - Lecturer
6 - No Rank
7 - All Instructional Staff

```{r setup, cache=TRUE, warning=FALSE}

# Load Packages
pacman::p_load(tidyverse, broom, janitor, readr, stargazer, skimr, tidymodels, magrittr,doParallel, fixest, modelsummary, kableExtra, did, gt, beepr, ggplot2, maps, synthdid, devtools, augsynth)

# Set directory
setwd("C:/Working_Projects/HigherEdLabor")

# Load Data
union <- read.csv("Written Data/union.csv")
pay <- read.csv("C:/Working_Projects/Research-Data-Storage/Data/IPEDS/combined_table.csv")

beep(1)
print("Setup Complete")

```

```{r data_cleaning, cache=TRUE, warning=FALSE}

# pay (combined_table) has one row per institution per year, union has multiple rows per institution per year (one for each union type)

# Join the two data frames, keeping all rows from `pay` and matching multiple rows from `union`
merged_data <- pay %>%
  left_join(union, by = "UNITID")


merged_data <- clean_names(merged_data, case = "lower_camel")


merged_data <- merged_data %>% 
  mutate(
    after_recog_date = ifelse(yr >= recognitionDate, 1, 0), # Indicator for if the union is recognized
    after_cba_date = ifelse(yr >= cbaDate, 1, 0), # Indicator for reached collective bargaining agreement
    after_recog_date = ifelse(recognitionDate %>% is.na(), 0, after_recog_date), 
    after_cba_date = ifelse(cbaDate %>% is.na(), 0, after_cba_date),
    endowment = coalesce(f1AEndowment, f2Endowment),# Combine the f1, f2, f3 predictors
    totalRevenue = coalesce(f1ATotalRevenue, f2TotalRevenue),
    totalGrants = coalesce(f1ATotalGrants, f2TotalGrants),
    expenses = coalesce(f1AExpenses, f2Expenses, f3Expenses),
    tuitionAndFees = coalesce(f1ATuitionAndFees, f2TuitionAndFees),
    totalAssets = coalesce(f1ATotalAssets, f2TotalAssets, f3TotalAssets),
    f1a_tuition_percap = f1ATuitionAndFees / ugTot, # Create a measure of tuition per student
    f1a_grants_percap = f1ATotalGrants / ugTot,
    f2_tuition_percap = f2TuitionAndFees / ugTot,
    f2_grants_percap = f2TotalGrants / ugTot,
    ft = ifelse(grepl("FT", unitComposition), 1, 0),# Add indicator for type of unit composition
    pt = ifelse(grepl("PT", unitComposition), 1, 0),
    ttt = ifelse(grepl("TTT", unitComposition), 1, 0),
    ntt = ifelse(grepl("NTT", unitComposition), 1, 0)

  # The rest will be done after the mdf transform to save compute time
    
  )

# Clean data frame so only 1 row per inst. year.
# Need different CBA dates for different union comps at same school

indicators <- c(
  "ft", "pt", "ttt", "ntt", "af_cba_n", "af_cba_t", "cba_date_t", "cba_date_n", "carnegie", "sector", "hasMedSchool", "hbcu",
  "tribal", "hasVet", "hasLaw", "hasEngineering", "hasDentistry", "hasBusiness"
)

faculty <- c(
  "instTotTR1", "instTotTR2", "instTotTR3", "instTotTR4", "instTotTR5", "instTotTR6", "instTotTR7",
  "instFtTR1", "instFtTR2", "instFtTR3", "instFtTR4", "instFtTR5", "instFtTR6", "instFtTR7",
  "instTotMR1", "instTotMR2", "instTotMR3", "instTotMR4", "instTotMR5", "instTotMR6", "instTotMR7",
  "instFtMR1", "instFtMR2", "instFtMR3", "instFtMR4", "instFtMR5", "instFtMR6", "instFtMR7",
  "instTotWR1", "instTotWR2", "instTotWR3", "instTotWR4", "instTotWR5", "instTotWR6", "instTotWR7",
  "instFtWR1", "instFtWR2", "instFtWR3", "instFtWR4", "instFtWR5", "instFtWR6", "instFtWR7",
  "avgSal9MoeqTR1", "avgSal9MoeqTR2", "avgSal9MoeqTR3", "avgSal9MoeqTR4", "avgSal9MoeqTR5", "avgSal9MoeqTR6", "avgSal9MoeqTR7",
  "avgSal9MoeqMR1", "avgSal9MoeqMR2", "avgSal9MoeqMR3", "avgSal9MoeqMR4", "avgSal9MoeqMR5", "avgSal9MoeqMR6", "avgSal9MoeqMR7",
  "avgSal9MoeqWR1", "avgSal9MoeqWR2", "avgSal9MoeqWR3", "avgSal9MoeqWR4", "avgSal9MoeqWR5", "avgSal9MoeqWR6", "avgSal9MoeqWR7"
)

keepers <- c("institutionName", "stateCode", "city", "x") # Removed "unitid", "yr", 

nonind <- setdiff(names(merged_data), c(indicators, "unitid", "yr")) # removed keepers and faculty

# This MIGHT BE WRONG.
# Gets rid of duplicate rows by summing things like different contract types 
# ^ This should be irrelevant since we expect combined_table to have one row per inst year
# and takes the max of any indicator in a yr
#
# Currently, merged_data is 1,000,000 lines, which is way more than we expect (in the 10,000s per year, a few hundred thousand total)
# This is using all my RAM to look at, and means that more data cleaning is needed.
# ^ Actually, MDF CLEAN is a reasonable length but with some missing data (2005-2006)
#
# Currently, the 2005-06 Access database has 7000 lines in HD, 37000 in SAL, and 445000 in the output table
# It appears that this problem only exists for the years where i removed CONTRACT from SQL, 2005 and 2006
#
# After picking just one contract type, the data looks reasonable in length
mdf_clean <- merged_data %>% 
  mutate(af_cba_t = ifelse(after_cba_date==1 & ttt==1, 1, 0),
         af_cba_n = ifelse(after_cba_date==1 & ntt==1, 1, 0),
         # add ttt and ntt specific date columns for did
         cba_date_t = ifelse(ttt==1, cbaDate, NA),
         cba_date_n = ifelse(ntt==1, cbaDate, NA)
         ) %>% 
  group_by(unitid, yr) %>% 
  summarise(across(all_of(nonind), first),
            #across(all_of(faculty), sum),
            #across(all_of(keepers), first), 
            across(all_of(indicators), max),
            .groups = "drop")
  
mdf_clean %<>% 
  mutate(
    logsal1 = log(avgSal9MoeqTR1),
    logsal2 = log(avgSal9MoeqTR2),
    logsal3 = log(avgSal9MoeqTR3),
    logsal4 = log(avgSal9MoeqTR4),
    logsal5 = log(avgSal9MoeqTR5),
    logsal6 = log(avgSal9MoeqTR6),
    logsal7 = log(avgSal9MoeqTR7)
  ) %>% 
  group_by(unitid) %>% 
  mutate(cba_ever = ifelse(any(after_cba_date == 1), 1, 0)) %>% 
  ungroup() %>% 
  mutate(t = yr - cbaDate, 

    # Divide all avg sal measures by 9 for 2011 and earlier to correct for mistake in SQL code
    # Instead of multiplying by 9 to get avg sal for post 2011, divide total salary outlays by # of full time workers
    # Lets actually just abandon the entire 9moeq and do everything by the above method
    # 2015 and earlier, instTot is not reported, instead use isntFt
    avgSalTR1 = ifelse(yr <= 2015, totSalTR1 / instFtTR1,  totSalTR1 / instTotTR1), 
    avgSalTR2 = ifelse(yr <= 2015, totSalTR2 / instFtTR2,  totSalTR2 / instTotTR2), 
    avgSalTR3 = ifelse(yr <= 2015, totSalTR3 / instFtTR3,  totSalTR3 / instTotTR3), 
    avgSalTR4 = ifelse(yr <= 2015, totSalTR4 / instFtTR4,  totSalTR4 / instTotTR4), 
    avgSalTR5 = ifelse(yr <= 2015, totSalTR5 / instFtTR5,  totSalTR5 / instTotTR5), 
    avgSalTR6 = ifelse(yr <= 2015, totSalTR6 / instFtTR6,  totSalTR6 / instTotTR6), 
    avgSalTR7 = ifelse(yr <= 2015, totSalTR7 / instFtTR7,  totSalTR7 / instTotTR7), 

    avgSalMR1 = ifelse(yr <= 2015, totSalMR1 / instFtMR1,  totSalMR1 / instTotMR1), 
    avgSalMR2 = ifelse(yr <= 2015, totSalMR2 / instFtMR2,  totSalMR2 / instTotMR2), 
    avgSalMR3 = ifelse(yr <= 2015, totSalMR3 / instFtMR3,  totSalMR3 / instTotMR3), 
    avgSalMR4 = ifelse(yr <= 2015, totSalMR4 / instFtMR4,  totSalMR4 / instTotMR4), 
    avgSalMR5 = ifelse(yr <= 2015, totSalMR5 / instFtMR5,  totSalMR5 / instTotMR5), 
    avgSalMR6 = ifelse(yr <= 2015, totSalMR6 / instFtMR6,  totSalMR6 / instTotMR6), 
    avgSalMR7 = ifelse(yr <= 2015, totSalMR7 / instFtMR7,  totSalMR7 / instTotMR7), 

    avgSalWR1 = ifelse(yr <= 2015, totSalWR1 / instFtWR1,  totSalWR1 / instTotWR1), 
    avgSalWR2 = ifelse(yr <= 2015, totSalWR2 / instFtWR2,  totSalWR2 / instTotWR2), 
    avgSalWR3 = ifelse(yr <= 2015, totSalWR3 / instFtWR3,  totSalWR3 / instTotWR3), 
    avgSalWR4 = ifelse(yr <= 2015, totSalWR4 / instFtWR4,  totSalWR4 / instTotWR4), 
    avgSalWR5 = ifelse(yr <= 2015, totSalWR5 / instFtWR5,  totSalWR5 / instTotWR5), 
    avgSalWR6 = ifelse(yr <= 2015, totSalWR6 / instFtWR6,  totSalWR6 / instTotWR6), 
    avgSalWR7 = ifelse(yr <= 2015, totSalWR7 / instFtWR7,  totSalWR7 / instTotWR7)
  )

  mdf_clean %<>% 
    mutate(
      logsal1 = log(avgSalTR1),
      logsal2 = log(avgSalTR2),
      logsal3 = log(avgSalTR3),
      logsal4 = log(avgSalTR4),
      logsal5 = log(avgSalTR5),
      logsal6 = log(avgSalTR6),
      logsal7 = log(avgSalTR7), 
      logTuition = ifelse(log(tuitionAndFees) == -Inf, NA, log(tuitionAndFees))
    )

  beep(1) # Tell me when its done
  print("Data Cleaning Complete")
  

  # FINDING
  # post-2011 use FXInstructionalSalariesWages for total salary outlays
  # post-2011 use tenuredFaculty +nonTenuredFaculty + tenureTrackFaculty for total faculty count

  write.csv(mdf_clean, "Written Data/mdf_clean.csv", row.names = FALSE)


```



```{r data_vis, cache=TRUE, warning=FALSE, fig.width=10, fig.height=5}



# plot avg salaries to see if they look reasonable

## Data Prep
#--------------------------------------------

mdf_plot <- mdf_clean %>% 
  mutate(
    instTotTR1 = ifelse(yr <= 2015, instFtTR1, instTotTR1)
  ) %>%
  group_by(yr) %>% 
  summarise(
    avgSal1 = mean(avgSalTR1, na.rm = TRUE),
    avgSal2 = mean(avgSalTR2, na.rm = TRUE),
    avgSal3 = mean(avgSalTR3, na.rm = TRUE),
    avgSal4 = mean(avgSalTR4, na.rm = TRUE),
    avgSal5 = mean(avgSalTR5, na.rm = TRUE),
    avgSal6 = mean(avgSalTR6, na.rm = TRUE),
    avgSal7 = mean(avgSalTR7, na.rm = TRUE), 

    totSal1 = mean(totSalTR1, na.rm =TRUE), 
    instTot = mean(instTotTR1, na.rm=TRUE),
    instFt = mean(instFtTR1, na.rm = TRUE), 
    ft = mean(instFtTR1, na.rm =TRUE), 
    tot = mean(instTotTR1, na.rm=TRUE)
  ) %>%
  pivot_longer(cols = starts_with("avgSal"), names_to = "Rank", values_to = "AvgSalary")


## Plot
#--------------------------------------------

ggplot(mdf_plot, aes(x = yr, y = AvgSalary, color = Rank)) + # I dont understand why this is fucked, it doesnt look like the data??
  geom_point() +
  geom_line() +
  labs(color = "Rank",
       title = "Average Salary by Rank",
       y = "Average Salary",
       x = "Year") +
  geom_vline(xintercept = 2011) 

# There is a shift after 2011, which is when the data reporting shifts
# Remember that pre 2012, salaries are annual, and after salaries are monthly * 9. What if not everyone is on a 9 month contract?

ggplot(mdf_plot, aes(x = yr, y = totSal1)) + # This one looks normal # NOT since removing CONTRACT from SQL
geom_point()+
geom_line() +
labs(title = "Average Total Salary Expenditure for All Faculty", 
     y = "Total Salary Outlays (From totSalTR1)",
     x = "Year"
)

ggplot(mdf_plot, aes(x = yr, y = instTot)) + # Why is the avg num of staff so low between 2005-2011? In testdf, instTot is NA through 2015
geom_point()+
geom_line()+
labs(title = "Average Total Number of Faculty", 
     y = "Total Number of Faculty (From instTotTR1)",
     x = "Year"
)

ggplot(mdf_plot, aes(x = yr, y = ft)) + # Why is this the same as the above one??
geom_point()+
geom_line()+
labs(title = "Average Total Number of Faculty", 
     y = "Total Number of Faculty (From instFtTR1)",
     x = "Year"
)
# its because instTot isnt pulling from the right access table for those years (2005-2011)

# 2015 and earlier, instTot is not reported

ggplot(mdf_plot %>% filter(Rank == "avgSal1"), aes(x = yr, y = AvgSalary)) +
geom_point()+
geom_line() +
labs(title = "Average Salary by Dividing Total Salary by Total Faculty", 
     y = "Average Salary",
     x = "Year"
) # Jump after 2011 because data is different! This looks good but its NOT


# What if we make an indicator for "post2011" instead of trying to fix the data?




# Try plotting just one school
mdf_solo <- mdf_clean %>% 
  filter(institutionName == "Harvard University")


ggplot(mdf_solo, aes(x = yr, y = avgSalTR7)) + 
  geom_point()+
  geom_line() +
  labs(title = "Harvard University Average Salary All Profs", 
       y = "Average Salary",
       x = "Year"
  )

ggplot(mdf_solo, aes(x = yr, y = instFtTR7)) + 
  geom_point()+
  geom_line() +
  labs(title = "Harvard University Full-Time Faculty Count", 
       y = "Full-Time Faculty Count",
       x = "Year"
  )

ggplot(mdf_solo, aes(x = yr, y = totSalTR7)) + 
  geom_point()+
  geom_line() +
  labs(title = "Harvard University Total Salary All Profs", 
       y = "Total Salary",
       x = "Year"
  )


  # Plot share of units with af_cba_t == 1 over time
  mdf_share <- mdf_clean %>% 
    group_by(yr) %>%
    summarise(
      share_af_cba_t = mean(af_cba_t, na.rm = TRUE)
    )

  ggplot(mdf_share, aes(x = yr, y = share_af_cba_t)) +
    geom_line() +
    geom_point() +
    labs(
      title = "Share of Units with Tenure-Track CBA Over Time",
      x = "Year",
      y = "Share (af_cba_t == 1)"
    ) +
    scale_y_continuous(limits = c(0, 1)) +
    theme_minimal()

    test <- mdf_clean %>% filter(institutionName == "University of Oregon") %>% select(cbaDate, institutionName, af_cba_t, af_cba_n, yr)

    # General Sense of Data

  # Treatment Year
    ggplot(mdf_clean, aes(x = cbaDate)) +
    geom_histogram(binwidth = 1, fill = "#444444", color = "black") +
    labs(
      title = "Distribution of CBA Treatment Years",
      x = "CBA Date (Year)",
      y = "Frequency"
    ) +
    theme_minimal() + theme(aspect.ratio = 9/16)

    # Treatment Share Over Time
    mdf_plot_1 <- mdf_clean %>% 
      group_by(yr) %>% 
      summarise(
        mean_af_cba_t = mean(af_cba_t, na.rm = TRUE),
        mean_af_cba_n = mean(af_cba_n, na.rm = TRUE)
      )

    ggplot(mdf_plot_1, aes(x = yr)) +
      geom_line(aes(y = mean_af_cba_t, color = "Tenure-Track"), linewidth = 1) +
      geom_line(aes(y = mean_af_cba_n, color = "Non-Tenure-Track"), linewidth = 1) +
      labs(
        title = "Share of Units with CBA Treatment Over Time",
        x = "Year",
        y = "Share (CBA Treatment)",
        color = "CBA Type"
      ) +
      scale_color_manual(values = c("Tenure-Track" = "#444444", "Non-Tenure-Track" = "#aaaaaa")) +
      theme_minimal() + 
      theme(aspect.ratio = 9/16) + 
      scale_y_continuous(labels = scales::label_percent())


    # Plot average salary over time for each rank
    mdf_plot_2 <- mdf_clean %>% 
      group_by(yr) %>% 
      summarise(
        avgSalTR1 = mean(avgSalTR1, na.rm = TRUE) / 1e3, # Scale to thousands for better visualization
        avgSalTR2 = mean(avgSalTR2, na.rm = TRUE) / 1e3,
        avgSalTR3 = mean(avgSalTR3, na.rm = TRUE) / 1e3,
        avgSalTR4 = mean(avgSalTR4, na.rm = TRUE) / 1e3,
        avgSalTR5 = mean(avgSalTR5, na.rm = TRUE) / 1e3,
        avgSalTR6 = mean(avgSalTR6, na.rm = TRUE) / 1e3,
        avgSalTR7 = mean(avgSalTR7, na.rm = TRUE) / 1e3
      ) %>% 
      pivot_longer(cols = starts_with("avgSal"), names_to = "Rank", values_to = "AvgSalary")

   ggplot(mdf_plot_2, aes(x = yr, y = AvgSalary, color = factor(Rank), linetype = factor(Rank))) +
  geom_line(linewidth = 1) + 
  labs(
    title = "Average Salary by Rank Over Time",
    x = "Year",
    y = "Average Salary (Thousands of Dollars)",
    color = "Rank",
    linetype = "Rank"
  ) +
  scale_linetype_manual(
    values = c("avgSalTR1" = "solid", "avgSalTR2" = "dashed", "avgSalTR3" = "dotted", 
               "avgSalTR4" = "dotdash", "avgSalTR5" = "longdash", "avgSalTR6" = "twodash"),
    labels = c("avgSalTR1" = "Professors", "avgSalTR2" = "Associate Prof", "avgSalTR3" = "Assistant Prof",
               "avgSalTR4" = "Instructor", "avgSalTR5" = "Lecturer", "avgSalTR6" = "All Instructional Staff")
  ) +
  scale_color_manual(
    values = c("avgSalTR1" = "#222222", "avgSalTR2" = "#444444", "avgSalTR3" = "#666666", 
               "avgSalTR4" = "#888888", "avgSalTR5" = "#aaaaaa", "avgSalTR6" = "#cccccc"),
    labels = c("avgSalTR1" = "Professors", "avgSalTR2" = "Associate Prof", "avgSalTR3" = "Assistant Prof",
               "avgSalTR4" = "Instructor", "avgSalTR5" = "Lecturer", "avgSalTR6" = "All Instructional Staff")
  ) +
  theme_minimal() +
  theme(aspect.ratio = 9/16)


  # Salary range histogram for each schools average over last 10 years
  mdf_plot_3 <- mdf_clean %>% 
    filter(yr >= 2010, ugTot > 500) %>% 
    group_by(institutionName) %>% 
    summarise(
      avgSalTR1 = mean(avgSalTR1, na.rm = TRUE) / 1e3, # Scale to thousands for better visualization
      avgSalTR2 = mean(avgSalTR2, na.rm = TRUE) / 1e3,
      avgSalTR3 = mean(avgSalTR3, na.rm = TRUE) / 1e3,
      avgSalTR4 = mean(avgSalTR4, na.rm = TRUE) / 1e3,
      avgSalTR5 = mean(avgSalTR5, na.rm = TRUE) / 1e3,
      avgSalTR6 = mean(avgSalTR6, na.rm = TRUE) / 1e3,
      avgSalTR7 = mean(avgSalTR7, na.rm = TRUE) / 1e3
    )

  ggplot(mdf_plot_3, aes(x = avgSalTR7)) +
    geom_histogram(position = "identity", bins = 30) +
    labs(
      title = "Distribution of Average Salaries(2010-2020)",
      x = "Average Salary (Thousands of Dollars)",
      y = "Count of Institutions",
      fill = "Rank"
    ) +
    theme_minimal() +
    theme(aspect.ratio = 9/16)

    # Horizontal bar chart of total students (ugTot) by state

  mdf_plot_4 <- mdf_clean %>% 
    group_by(stateCode) %>% 
    summarise(totalStudents = sum(ugTot, na.rm = TRUE) / 1e6) %>% 
    filter(totalStudents > 0.5) %>%
    arrange(totalStudents)

  ggplot(mdf_plot_4, aes(x = totalStudents, y = reorder(stateCode, totalStudents))) +
    geom_bar(stat = "identity", fill = "#444444") +
    labs(
      title = "Total Undergraduate Students by State",
      x = "Total Undergraduate Students (Millions)",
      y = "State"
    ) +
    theme_minimal() +
    theme(aspect.ratio = 12/16) + theme(axis.text.y = element_text(size = 13))

```{r test_regression, cache=TRUE, warning=FALSE}
# Lets see if the data is statistically different after 2011

mdf_reg <- mdf_clean %>%
  mutate(post2011 = ifelse(yr > 2011, 1, 0))

models <- lapply(c("avgSalTR1", "avgSalTR2", "avgSalTR3", "avgSalTR4", "avgSalTR5", "avgSalTR6", "avgSalTR7"), function(outcome) {
  feols(as.formula(paste("log(", outcome, ") ~ post2011")), data = mdf_reg, vcov = "hetero")
})

modelsummary(
  models,
  title = "Post-2011 Treatment Effect",
  output = "gt",
  shape = term ~ model,
  stars = TRUE
)

# Significance here tells us that there is likely something different about the data after 2011
# I can't find anything in post-2011 data that explains this

# Also, when i change the year for this indicator, the results do not change

```
## Balance Table
```{r balance_table, cache=TRUE, warning=FALSE}
pre_survey_vars <- c(
  "f1AEndowment",
  "f2Endowment",
  "f2TotalAssets",
  "f1ATotalAssets",
  "applicantsTot",
  "instTotTR1",
  "instTotTR2",
  "instTotTR3",
  "instTotTR4",
  "instTotTR5",
  "instTotTR6",
  "instTotTR7",
  "avgSalTR1"
)


models <- lapply(pre_survey_vars, function(var) {
  feols(as.formula(paste("log(",var, ")~ cba_ever")), data = mdf_clean , vcov = "hetero")
})
names(models) <- pre_survey_vars

modelsummary(
  models,
  title = "Balance Table (in logs)",
  output = "gt",  # changed to gt for compatibility
  shape = model ~ term,
  stars = TRUE
)




```
 ## Pre-Trends Plots
```{r pre_trends, cache=TRUE, warning=FALSE}

outcomes <- c("avgSalTR1", "avgSalTR2", "avgSalTR3", "avgSalTR4", "avgSalTR5", "avgSalTR6", "avgSalTR7" )

models_pt <- lapply(outcomes, function(out) {
  feols(
    as.formula(paste0("log(",out, ")~ i(t, ref = 0) | unitid + yr")), data = mdf_clean
  )
})

modelsummary(
  models_pt,
  title = "Pre-Trends Test",
  output = "gt",
  shape = term~ model,
  stars = TRUE,
  coef_map = c(
    "t::-1" = "1 Year Before Treatment",
    "t::-2" = "2 Years Before Treatment",
    "t::-3" = "3 Years Before Treatment",
    "t::-4" = "4 Years Before Treatment",
    "t::-5" = "5 Years Before Treatment"
  )
)

# Summary Statistics
lapply(outcomes, function(out) {
  mdf_clean %>%
    summarise(
      mean = mean(get(out), na.rm = TRUE),
      median = median(get(out), na.rm = TRUE),
      sd = sd(get(out), na.rm = TRUE),
      p10 = quantile(get(out), 0.1, na.rm = TRUE),
      p90 = quantile(get(out), 0.9, na.rm = TRUE)
    )
})

# Plot
mdf_sum <- mdf_clean %>% 
  filter(!(after_cba_date)) %>% 
  group_by(cba_ever, yr) %>%
  summarise(
    tr1 = mean(avgSalTR1, na.rm = TRUE),
    tr2 = mean(avgSalTR2, na.rm = TRUE),
    tr3 = mean(avgSalTR3, na.rm = TRUE),
    tr4 = mean(avgSalTR4, na.rm = TRUE),
    tr5 = mean(avgSalTR5, na.rm = TRUE),
    tr6 = mean(avgSalTR6, na.rm = TRUE),
    tr7 = mean(avgSalTR7, na.rm = TRUE)
  ) 


plotsout <- lapply(c("tr1", "tr2", "tr3", "tr4", "tr5", "tr6", "tr7"), function(out) {
  ggplot(mdf_sum,aes(x = yr))+
  geom_smooth(aes(y = log(!!sym(out)), color = factor(cba_ever)))
})

# Don't forget that this puts out several plots


plotsout[1]
plotsout[2]
plotsout[3]
plotsout[4]
plotsout[5]
plotsout[6]
plotsout[7]

```

```{r did, cache=TRUE, warning=FALSE, }
dep_vars <- c(
  "logsal1", 
  "logsal2", 
  "logsal3", 
  "logsal4", 
  "logsal5", 
  "logsal6", 
  "logsal7"
  )


plotsout <- lapply(dep_vars[1:3], function(dep){
  ggdid(aggte(att_gt(yname = dep,
                     tname = "yr",
                     idname = "unitid",
                     gname = "cba_date_t", 
                     xformla = ~1,
                     est_method = "reg",
                     control_group = "notyettreated",
                     data = mdf_clean), type = "dynamic", na.rm=TRUE), title = dep)

})

plotsout[4:6] <- lapply(dep_vars[4:6], function(dep){
  ggdid(aggte(att_gt(yname = dep,
                     tname = "yr",
                     idname = "unitid",
                     gname = "cba_date_n",
                     xformla = ~1,
                     est_method = "reg",
                     control_group = "notyettreated",
                     data = mdf_clean), type = "dynamic", na.rm=TRUE), title = dep)

})

plotsout[7]<- lapply(dep_vars[7], function(dep){
  ggdid(aggte(att_gt(yname = dep,
                     tname = "yr",
                     idname = "unitid",
                     gname = "cba_date_n", 
                     xformla = ~1,
                     est_method = "reg",
                     control_group = "notyettreated",
                     data = mdf_clean), type = "dynamic", na.rm=TRUE), title = dep)

})

# Get plots for tuition as well

plotsout[[8]] <- ggdid(aggte(att_gt(yname = "logTuition",
                     tname = "yr",
                     idname = "unitid",
                     gname = "cba_date_n", 
                     xformla = ~1,
                     est_method = "reg",
                     control_group = "notyettreated",
                     data = mdf_clean), type = "dynamic", na.rm=TRUE), title = "Tuition and Fees") + labs(title = "Tuition Effects from NT") + theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
 )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-16, 16, by = 4))

plotsout[[9]] <- ggdid(aggte(att_gt(yname = "logTuition",
                     tname = "yr",
                     idname = "unitid",
                     gname = "cba_date_t", 
                     xformla = ~1,
                     est_method = "reg",
                     control_group = "notyettreated",
                     data = mdf_clean), type = "dynamic", na.rm=TRUE), title = "Tuition and Fees") + labs(title = "Tuition Effects from TT") + theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
 )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-16, 16, by = 4))

titles <- c(
  "Full Professors",
  "Associate Professors",
  "Assistant Professors",
  "Instructors",
  "Lecturers",
  "No Rank",
  "All Instructional Staff",
  "Tuition Effects from NT", 
  "Tuition Effects from TT"
)
  

for(p in 1:length(titles)){

  df = plotsout[[p]]$data

  df$lower <- df$att - 1.96 * df$att.se
  df$upper <- df$att + 1.96 * df$att.se

  # Layer 1 = error bars
  plotsout[[p]]$layers[[1]]$aes_params$linewidth <- 3   # thicker error bars
  plotsout[[p]]$layers[[1]]$aes_params$width = 0  # change color

  # Layer 2 = points
  plotsout[[p]]$layers[[2]]$aes_params$size <- 1         # larger points


  ggsave(filename = paste0("plot_", titles[p], ".png"),
    plot = plotsout[[p]]+ labs(title = titles[p], subtitle = paste0("Mean Estimate: ", round(mean(df$att), 4), " Mean SE : ", round(mean(df$att.se), 4))) + theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    plot.subtitle = element_text(
      hjust = 0.5,
      size = 16,
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
  )+ theme(text = element_text(size = 40), plot.title = element_text(size = 45), plot.subtitle = element_text(size = 35)) +
  scale_x_continuous(breaks = seq(-16, 16, by = 4))
  ,
   width = 12, height = 6, dpi = 300
  ) 

}
# Use ggsave or aspect_ratio

plotsout[[1]] + labs(title = "Full Professors") + theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
  )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-16, 16, by = 4))

plotsout[[2]] + labs(title = "Associate Professors") + theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
 )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-16, 16, by = 4))


plotsout[[3]] + labs(title = "Assistant Professors") + theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
 )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-16, 16, by = 4))

  
plotsout[[4]] + labs(title = "Instructors") + theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
 )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-16, 16, by = 4))

  
plotsout[[5]] + labs(title = "Lecturers") + theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
 )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-16, 16, by = 4))

plotsout[[6]] + labs(title = "No Rank") + theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
 )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-16, 16, by = 4))

  
plotsout[[7]] + labs(title = "All Instructional Staff") + theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
 )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-16, 16, by = 4))

  


plotsout[[8]]

```

```{r monteCarlo, cache=TRUE, warning=FALSE}

# Generate data for synthetic did 

years <- 2005:2024
t_years <- 2010:2020
len <- length(years)
t_len <- length(t_years)
n = 500
tenure_actual = -0.1
non_tenure_actual = -0.2
prop_treated = 0.3

runsim <- function(years, t_years, n, tenure_actual, non_tenure_actual, prop_treated, out = TRUE, b = TRUE, std = 0.1) {

mcdf <- data.frame(
  unitid = rep(1:n, each = len),
  yr = rep(years, times = n)
) %>% group_by(unitid) %>% 
  mutate(
    cba_date_t = sample(c(NA, t_years), size = 1, replace = TRUE, prob = c(prop_treated, rep((1 - prop_treated) / t_len, t_len))),
    cba_date_n = sample(c(NA, t_years), size = 1, replace = TRUE, prob = c(prop_treated, rep((1 - prop_treated) / t_len, t_len))),
    af_cba_t = ifelse(!is.na(cba_date_t) & yr >= cba_date_t, 1, 0),
    af_cba_n = ifelse(!is.na(cba_date_n) & yr >= cba_date_n, 1, 0),
    logsal1 = rnorm(n(), mean = 10 + af_cba_t * tenure_actual + af_cba_n * non_tenure_actual + (yr - min(years)) * 0.02, sd = std),
    always_treated = ifelse(all(af_cba_t == 1) | all(af_cba_n == 1), 1, 0)
  )%>% 
  ungroup() %>% 
  filter(always_treated == 0)


  
  synthmod_t <- multisynth(logsal1 ~ af_cba_t, unitid, yr, data = mcdf, fixedeff = TRUE, time_cohort = TRUE)
  #beep(1)
  out_t <- synthmod_t %>% summary()
  #plot(out_t)
  #beep(1)

  att_t <- out_t$att$Estimate[out_t$att$Level == "Average" & is.na(out_t$att$Time)]

  synthmod_n <- multisynth(logsal1 ~ af_cba_n, unitid, yr, data = mcdf, fixedeff = TRUE, time_cohort = TRUE)
  #beep(1)
  out_n <- synthmod_n %>% summary()
  #plot(out_n)
  #beep(1)

    att_n <- out_n$att$Estimate[out_n$att$Level == "Average" & is.na(out_n$att$Time)]

error_t <- round((att_t - tenure_actual) / tenure_actual, 2)
error_n <- round((att_n - non_tenure_actual) / non_tenure_actual, 2)

if(out){
cat("---------------------------------\n","Monte Carlo Simulation Results\n","--------------------------------\n")
cat("N = ", n, "\n")
cat("Total Years: ", len, "\n")
cat("Treatment Years: ", t_len, "\n")
cat("\nTenure effect estimate:", att_t, "\n")
cat("Error of " , error_t, "%\n")
cat("\nNon-tenure effect estimate:", att_n, "\n")
cat("Error of " , error_n, "%\n")
}
if(b){beep(1)}


out =  c(att_t, att_n, error_t, error_n)
return(out)
}

runsim(years, t_years, n, tenure_actual, non_tenure_actual, prop_treated)

## Test effect of varying n
nset <- 10^seq(1, 2, by = 0.5)

count <- 0
vary_n <- lapply(nset, function(n) {
  # run simulation
  res <- runsim(years, t_years, n, tenure_actual, non_tenure_actual, prop_treated, out = FALSE, b = FALSE)
  
  # update progress
  count <<- count + 1   # use <<- so it updates the outer variable
  pct <- round(count / length(nset) * 100, 2)
  cat("Progress: ", pct, "%\n")
  
  # return result
  res
})

# combine results into a dataframe
vndf <- do.call(rbind, vary_n) %>% 
  as.data.frame() %>% 
  rename(att_t = V1, att_n = V2, error_t = V3, error_n = V4) %>% 
  mutate(n = nset)

# regressions
feols((error_t)^2 ~ n, data = vndf, vcov = "hetero") %>% summary()
feols((error_n)^2 ~ n, data = vndf, vcov = "hetero") %>% summary()

## Test effect of varying n vs varying prop treated


quiet_runsim_safe <- function(n, prop_treated, years, t_years,
                              tenure_actual, non_tenure_actual,
                              std = 0.1) {
  len <- length(years)
  t_len <- length(t_years)
  
  # remove last year from t_years to ensure pre-treatment exists
  t_years_safe <- t_years[t_years < max(years)]
  
  mcdf <- data.frame(
    unitid = rep(1:n, each = len),
    yr = rep(years, times = n)
  )
  
  treat_dates <- data.frame(
    unitid = 1:n,
    cba_date_t = sample(c(NA, t_years_safe),
                        size = n, replace = TRUE,
                        prob = c(1 - prop_treated,
                                 rep(prop_treated / length(t_years_safe),
                                     length(t_years_safe)))),
    cba_date_n = sample(c(NA, t_years_safe),
                        size = n, replace = TRUE,
                        prob = c(1 - prop_treated,
                                 rep(prop_treated / length(t_years_safe),
                                     length(t_years_safe))))
  )
  
  # force at least one untreated unit per outcome
  if(all(!is.na(treat_dates$cba_date_t))) treat_dates$cba_date_t[1] <- NA
  if(all(!is.na(treat_dates$cba_date_n))) treat_dates$cba_date_n[1] <- NA
  
  mcdf <- mcdf %>%
    left_join(treat_dates, by = "unitid") %>%
    mutate(
      af_cba_t = ifelse(!is.na(cba_date_t) & yr >= cba_date_t, 1, 0),
      af_cba_n = ifelse(!is.na(cba_date_n) & yr >= cba_date_n, 1, 0),
      logsal1 = 10 + af_cba_t * tenure_actual + af_cba_n * non_tenure_actual +
                (yr - min(yr)) * 0.02 + rnorm(nrow(.), 0, std)
    )
  
  # safety check
  if(length(unique(mcdf$af_cba_t)) < 2 || length(unique(mcdf$af_cba_n)) < 2) {
    return(tibble(
      n = n,
      prop_treated = prop_treated,
      att_t = NA_real_,
      att_n = NA_real_,
      error_t = NA_real_,
      error_n = NA_real_
    ))
  }
  
  synthmod_t <- multisynth(logsal1 ~ af_cba_t, unitid, yr, data = mcdf, fixedeff = TRUE)
  synthmod_n <- multisynth(logsal1 ~ af_cba_n, unitid, yr, data = mcdf, fixedeff = TRUE)
  
  att_t <- summary(synthmod_t)$att$Estimate[summary(synthmod_t)$att$Level == "Overall"]
  att_n <- summary(synthmod_n)$att$Estimate[summary(synthmod_n)$att$Level == "Overall"]
  
  error_t <- round((att_t - tenure_actual)/tenure_actual, 2)
  error_n <- round((att_n - non_tenure_actual)/non_tenure_actual, 2)
  
  tibble(
    n = n,
    prop_treated = prop_treated,
    att_t = att_t,
    att_n = att_n,
    error_t = error_t,
    error_n = error_n
  )
}
# ------------------------------
# 2️⃣ Monte Carlo parameter grid
# ------------------------------

nset <- c(5000, 6000)
t_years <- 2010:2015         # sample sizes
prop_set <- c(0.1,0.2)   # proportion treated
reps <- 1                              # repetitions per combo

param_grid <- expand_grid(
  n = nset,
  prop_treated = prop_set,
  rep = 1:reps
)

# ------------------------------
# 3️⃣ Run parallel Monte Carlo
# ------------------------------

plan(multisession, workers = parallel::detectCores() - 1)

set.seed(123)

results <- future_pmap_dfr(
  list(param_grid$n, param_grid$prop_treated),
  function(n, p) {
    quiet_runsim_safe(n, p, years, t_years, tenure_actual, non_tenure_actual)
  },
  .options = furrr_options(seed = TRUE),
  .progress = TRUE
)

# Merge rep column if needed
results <- results %>% bind_cols(param_grid %>% select(rep))


quiet_runsim_safe(500, 0.2, years, t_years, tenure_actual, non_tenure_actual)

```

```{r mdfclean, cache=TRUE, warning=FALSE}

pacman::p_load(tidyverse, broom, janitor, readr, stargazer, skimr, tidymodels, magrittr,doParallel, fixest, modelsummary, kableExtra, did, gt, beepr, ggplot2, maps, synthdid, devtools, augsynth, future, furrr)
setwd("C:/Working_Projects/HigherEdLabor")

mdf_clean <- read.csv("Written Data/mdf_clean.csv")
beep(1)

```

```{r synthdid, cache=TRUE, warning=FALSE}
red = c("University of Oregon", "Oregon State University", "Portland State University", "University of Utah", "Utah State University", "Brigham Young University")

mdf_aug_1 %>%
  group_by(unitid) %>%
  summarise(
    pre_periods = sum(yr < min(yr[af_cba_t == 1], na.rm = TRUE))
  ) %>%
  summary()

  mdf_aug_1 %>%
  group_by(unitid) %>%
  summarise(
    ever_treated = any(af_cba_t == 1, na.rm = TRUE)
  ) %>%
  table()

bad_units <- c(133526,134510,192147,202480,202514,210775,237640,376224, 133553,133979,134945,136701,186618,192165,202523,202541,210809,237686,376321)
# Remove always treated

mdf_clean %>% 
  group_by(unitid) %>% 
  mutate(always_treated = ifelse(all(af_cba_t == 1) | all(af_cba_n == 1), 1, 0), 
    logsal = log(avgSalTR1)) %>% 
  ungroup() %>% 
  filter(
    always_treated == 0, cbaDate < 2019 | is.na(cbaDate), 
    cbaDate > 2005 | is.na(cbaDate),
    ugTot > 100) %>% 
  select(
    unitid, institutionName, yr, cbaDate, stateCode, af_cba_t, af_cba_n, logsal1, logsal2, logsal3, logsal4, logsal5, logsal6, logsal7, ugTot, instFtTR7
  )  -> mdf_aug

  mdf_aug_1 <- mdf_aug %>% filter((logsal1 != -Inf), !is.na(logsal1), logsal1 != "Invalid Number")
  mdf_aug_2 <- mdf_aug %>% filter((logsal2 != -Inf), !is.na(logsal2), logsal2 != "Invalid Number")
  mdf_aug_3 <- mdf_aug %>% filter((logsal3 != -Inf), !is.na(logsal3), logsal3 != "Invalid Number")
  mdf_aug_4 <- mdf_aug %>% filter((logsal4 != -Inf), !is.na(logsal4), logsal4 != "Invalid Number")
  mdf_aug_5 <- mdf_aug %>% filter((logsal5 != -Inf), !is.na(logsal5), logsal5 != "Invalid Number")
  mdf_aug_6 <- mdf_aug %>% filter((logsal6 != -Inf), !is.na(logsal6), logsal6 != "Invalid Number")
  mdf_aug_7 <- mdf_aug %>% filter((logsal7 != -Inf), !is.na(logsal7), logsal7 != "Invalid Number")

  mdf_aug %>% group_by(unitid) %>% summarise(
    ever_treated_t = any(af_cba_t == 1, na.rm = TRUE),
    ever_treated_n = any(af_cba_n == 1, na.rm = TRUE)
  ) %>% ungroup() %>% filter(ever_treated_t == TRUE | ever_treated_n == TRUE) %>% nrow()


  synthmod <- multisynth(logsal1 ~ af_cba_t, unitid, yr, data = mdf_aug_1, fixedeff = FALSE, time_cohort = FALSE) 
  plot(synthmod %>% summary())

  synthmod2 <- multisynth(logsal2 ~ af_cba_t , unitid, yr, data = mdf_aug_2, fixedeff = FALSE, time_cohort = FALSE) 
  plot(synthmod2 %>% summary())

  synthmod3 <- multisynth(logsal3 ~ af_cba_t , unitid, yr, data = mdf_aug_3, fixedeff = FALSE, time_cohort = FALSE) 
  plot(synthmod3 %>% summary())

  synthmod4 <- multisynth(logsal4 ~ af_cba_n , unitid, yr, data = mdf_aug_4, fixedeff = FALSE, time_cohort = FALSE) 
  plot(synthmod4 %>% summary())

  # 5 breaks idk why
  synthmod5 <- multisynth(logsal5 ~ af_cba_n , unitid, yr, data = mdf_aug_5, fixedeff = FALSE, time_cohort = FALSE) 
  plot(synthmod5 %>% summary())
  # 6 too
  synthmod6 <- multisynth(logsal6 ~ af_cba_n , unitid, yr, data = mdf_aug_6, fixedeff = FALSE, time_cohort = FALSE) 
  plot(synthmod6 %>% summary())

  synthmod7 <- multisynth(logsal7 ~ af_cba_t , unitid, yr, data = mdf_aug_7, fixedeff = FALSE, time_cohort = FALSE) 
  plot(synthmod7 %>% summary())

  synthmod %>% summary()

  plotsout_syn <- lapply(dep_vars[1:3], function(dep){ 
    synthmod <- multisynth(as.formula(paste0(dep, " ~ af_cba_t")), unitid, yr, data = mdf_aug, fixedeff = FALSE, time_cohort = TRUE)
    plot(synthmod %>% summary(), title = dep)
  })

  plotsout_syn[4:6] <- lapply(dep_vars[4:6], function(dep){ 
    synthmod <- multisynth(as.formula(paste0(dep, " ~ af_cba_n")), unitid, yr, data = mdf_aug, fixedeff = FALSE, time_cohort = TRUE)
    plot(synthmod %>% summary(), title = dep)
  })

  
```

```{r regressions, cache=TRUE, warning=FALSE}

mdf_clean %<>% 
mutate(af_2011 = ifelse(yr > 2011, 1, 0))
# Regressions for each type of union
tr1g <- feols(log(avgSalTR1) ~af_cba_t + af_2011 |  yr + unitid, data = mdf_clean ) #tr1g denotes total (m+f) group 1 (Profs) grouped by union type
tr2g <- feols(log(avgSalTR2) ~af_cba_t + af_2011 |  yr + unitid, data = mdf_clean )
tr3g <- feols(log(avgSalTR3) ~af_cba_t + af_2011 |  yr + unitid, data = mdf_clean )
tr4g <- feols(log(avgSalTR4) ~af_cba_n + af_2011 |  yr + unitid, data = mdf_clean )
tr5g <- feols(log(avgSalTR5) ~af_cba_n + af_2011 |  yr + unitid, data = mdf_clean )
tr6g <- feols(log(avgSalTR6) ~af_cba_n + af_2011 |  yr + unitid, data = mdf_clean )
tr7g <- feols(log(avgSalTR7) ~after_cba_date + af_2011|  yr + unitid, data = mdf_clean )

modelsummary(list("Professors" = tr1g,
                  "Associate Prof" = tr2g,
                  "Assistant prof" = tr3g,
                  "Instructor" = tr4g),
             output = "kableExtra",
             stars = TRUE,
             title = "tenure Track after CBA date")

modelsummary(list("Lecturer" = tr5g,
                  "No Rank"= tr6g,
                  "All Inst. Staff" = tr7g),
             output = "kableExtra",
             stars = TRUE,
             title = "Non tenure Track after CBA date")
```

```{r test, cache=TRUE, warning=FALSE}
testdf <- mdf_clean %>%
  filter(
    institutionName == "University of Oregon"
    #unitid == 100654
  ) %>% 
  select(
    avgSal9MoeqTR1, 
    yr, 
    unitid, 
    institutionName, 
    ugTot, 
    instTotTR7, 
    totSalTR7, 
    instFtTR7, 
    avgSalTR7
  )%>% 
  mutate ( 
    avgSal = totSalTR7 / instFtTR7
  )


testdf2 <- merged_data %>% 
  filter(
    #institutionName == "University of Oregon"
    unitid == 100654
  ) %>% 
  select(
    avgSal9MoeqTR1, 
    yr, 
    unitid, 
    institutionName, 
    ugTot, 
    instTotTR1
  ) 
testdf3 <- pay %>% 
  filter(
    #institutionName == "University of Oregon"
    UNITID == 100654
  ) %>% 
  select(
    YR, 
    UNITID, 
    inst_ft_m_r1, 
    inst_tot_t_r1
  )

# Make sure we have right number of observations

ggplot(mdf_clean %>% group_by(unitid) %>% summarize(count = n()), aes(x = count)) + 
  geom_histogram()
# looks perfect, nothing shows up more than the number of years in the data
```

## THESIS STUFF:

```{r thesis, cache=TRUE, warning=FALSE}

# Import csv of state legislation
# Add legislation as a factor indicator (4 types) to each row
# Regress endowment (may have to add from Access data) on legislation indicator with inst/yr fe
# Also do DiD analysis since it makes more sense here

thesis_treatment <- read.csv("../Thesis/transparency.csv")
df_stateecon <- read.csv("../Thesis/stateecondata.csv")

df_stateecon %<>%
  rename("1998" = "X1998",
         "1999" = "X1999",
         "2000" = "X2000",
         "2001" = "X2001",
         "2002" = "X2002",
         "2003" = "X2003",
         "2004" = "X2004",
         "2005" = "X2005",
         "2006" = "X2006",
         "2007" = "X2007",
         "2008" = "X2008",
         "2009" = "X2009",
         "2010" = "X2010",
         "2011" = "X2011",
         "2012" = "X2012",
         "2013" = "X2013",
         "2014" = "X2014",
         "2015" = "X2015",
         "2016" = "X2016",
         "2017" = "X2017",  
         "2018" = "X2018",
          "2019" = "X2019",
          "2020" = "X2020", 
          "2021" = "X2021", 
          "2022" = "X2022", 
          "2023" = "X2023", 
          "2024" = "X2024"
         )

df_stateecon_long <- df_stateecon %>%
  select(-LineCode) %>%
  filter(grepl("GDP|PCE|Personal Income" , Description, ignore.case = TRUE)) %>% 
  pivot_longer(
    cols = -c(GeoName, Description, GeoFips), 
    names_to = "yr", 
    values_to = "value"
  ) %>% 
  pivot_wider(
    names_from = Description, 
    values_from = value
  ) %>% 
  mutate(yr = as.numeric(yr))%>%
  rename(stateName = GeoName) %>% 
  mutate(
    stateCode = case_when(
      stateName == "District of Columbia" ~ "DC",
      stateName == "United States" ~ "US",
      stateName == "Alabama" ~ "AL",
      stateName == "Alaska" ~ "AK",
      stateName == "Arizona" ~ "AZ",
      stateName == "Arkansas" ~ "AR",
      stateName == "California" ~ "CA",
      stateName == "Colorado" ~ "CO",
      stateName == "Connecticut" ~ "CT",
      stateName == "Delaware" ~ "DE",
      stateName == "Florida" ~ "FL",
      stateName == "Georgia" ~ "GA",
      stateName == "Hawaii" ~ "HI",
      stateName == "Idaho" ~ "ID",
      stateName == "Illinois" ~ "IL",
      stateName == "Indiana" ~ "IN",
      stateName == "Iowa" ~ "IA",
      stateName == "Kansas" ~ "KS",
      stateName == "Kentucky" ~ "KY",
      stateName == "Louisiana" ~ "LA",
      stateName == "Maine" ~ "ME",
      stateName == "Maryland" ~ "MD",
      stateName == "Massachusetts" ~ "MA",
      stateName == "Michigan" ~ "MI",
      stateName == "Minnesota" ~ "MN",
      stateName == "Mississippi" ~ "MS",
      stateName == "Missouri" ~ "MO",
      stateName == "Montana" ~ "MT",
      stateName == "Nebraska" ~ "NE",
      stateName == "Nevada" ~ "NV",
      stateName == "New Hampshire" ~ "NH",
      stateName == "New Jersey" ~ "NJ",
      stateName == "New Mexico" ~ "NM",
      stateName == "New York" ~ "NY",
      stateName == "North Carolina" ~ "NC",
      stateName == "North Dakota" ~ "ND",
      stateName == "Ohio" ~ "OH",
      stateName == "Oklahoma" ~ "OK",
      stateName == "Oregon" ~ "OR",
      stateName == "Pennsylvania" ~ "PA",
      stateName == "Rhode Island" ~ "RI",
      stateName == "South Carolina" ~ "SC",
      stateName == "South Dakota" ~ "SD",
      stateName == "Tennessee" ~ "TN",
      stateName == "Texas" ~ "TX",
      stateName == "Utah" ~ "UT",
      stateName == "Vermont" ~ "VT",
      stateName == "Virginia" ~ "VA",
      stateName == "Washington" ~ "WA",
      stateName == "West Virginia" ~ "WV",
      stateName == "Wisconsin" ~ "WI",
      stateName == "Wyoming" ~ "WY"
    )
  )

thesis_treatment %<>%
  rename(stateCode = "scode")

mdf_th <- mdf_clean %>% 
  left_join(thesis_treatment, by = "stateCode")

mdf_th %<>% 
  left_join(df_stateecon_long, by = c("stateCode", "yr"))

mdf_th %<>% 
  mutate(
    explicit = ifelse(Transparency == "Explicit", 1, 0),
    Transparency = as.factor(Transparency) 
         )

mdf_th$Transparency <- relevel(mdf_th$Transparency, ref = "None")



mdf_th <- mdf_th %>%
  mutate(after_leg_year = ifelse(yr >= leg_year, 1, 0),
         after_leg_year = ifelse(leg_year %>% is.na(), 0, after_leg_year),
         after_leg_year_explicit = ifelse(after_leg_year == 1 & explicit == 1, 1, 0), 
         leg_year_explicit = ifelse(explicit == 1, leg_year, NA)
         ) %>%
  mutate(leg_ever = ifelse(is.na(leg_year), 0, 1), 
         leg_ever_explicit = ifelse(is.na(leg_year_explicit), 0, 1),
         leg_in_timeframe = ifelse(leg_ever == 1 & leg_year > 2004, 1, 0)) %>% 
  mutate(t = yr - leg_year) %>% 
  mutate( 
    f1AEndowment = f1AEndowment / 1e6, # Scale endowment to millions
    f2Endowment = f2Endowment / 1e6,
    log_f1AEndowment = ifelse(is.na(f1AEndowment) | f1AEndowment == 0, NA, log(f1AEndowment)),
    log_f2Endowment = ifelse(is.na(f2Endowment)| f2Endowment == 0, NA, log(f2Endowment))
  ) %>%  
  rename(
    real_GDP_2017usd = "  Real GDP (millions of chained 2017 dollars) 1", 
    real_personal_income_2017usd =  "  Real personal income (millions of constant (2017) dollars) 2",
    real_PCE_2017usd = "  Real PCE (millions of constant (2017) dollars) 3"
  ) 

mdf_th %<>%  
  mutate( 
    real_GDP_2017usd = as.numeric(real_GDP_2017usd),  
    real_personal_income_2017usd = as.numeric(real_personal_income_2017usd), 
    real_PCE_2017usd = as.numeric(real_PCE_2017usd)
  )


# Did 

log_dep_vars <- c(
  "log_f1AEndowment", 
  "log_f2Endowment"
  )

dep_vars <- c(
  "f1AEndowment", 
  "f2Endowment"
  )

plotsout <- lapply(log_dep_vars, function(dep){
  ggdid(aggte(att_gt(yname = dep,
                     tname = "yr",
                     idname = "unitid",
                     gname = "leg_year_explicit",
                     xformla = ~1,
                     est_method = "reg",
                     control_group = "notyettreated",
                     data = mdf_th), type = "dynamic", na.rm=TRUE), title = dep)

})

plotsout[[1]] +
  labs(title = "Public Institutions", 
        y = "Group Time ATT Estimate", 
        x = "Year Relative to Treatment") +
  theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold", size = 35),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
  )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-12, 10, by = 3))


plotsout[[2]] +
  labs(title = "Private Non-Profit", 
        y = "Group Time ATT Estimate", 
        x = "Year Relative to Treatment") +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      size = 18,
      face = "bold",
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold", size = 35),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
  )+ theme(text = element_text(size = 50), plot.title = element_text(size = 60)) +
  scale_x_continuous(breaks = seq(-12, 10, by = 3))

ggdid(aggte(att_gt(yname = "log_f1AEndowment",
                     tname = "yr",
                     idname = "unitid",
                     gname = "leg_year_explicit",
                     xformla = ~1,
                     est_method = "reg",
                     control_group = "notyettreated",
                     data = mdf_th), type = "dynamic", na.rm=TRUE), title = "Group-Time Difference-in-Differences Estimates for Public Institutions") 


ggdid(aggte(att_gt(yname = "log_f2Endowment",
                     tname = "yr",
                     idname = "unitid",
                     gname = "leg_year_explicit",
                     xformla = ~1,
                     est_method = "reg",
                     control_group = "notyettreated",
                     data = mdf_th), type = "dynamic", na.rm=TRUE), title = "Group-Time Difference-in-Differences Estimates for Private Non-Profit Institutions")

# Test pre-trends 

mdf_pre <- mdf_th %>% 
  filter(!(after_leg_year_explicit)) %>% 
  group_by(yr, leg_ever_explicit) %>% 
  summarise(f1AEndowment = mean(f1AEndowment, na.rm = TRUE), 
            f2Endowment = mean(f2Endowment, na.rm = TRUE)
            )
  


ggplot(data = mdf_pre, aes(x = yr, y = f1AEndowment, color = as.factor(leg_ever_explicit))) +
  geom_line()+ 
  scale_color_manual(
    values = c("red", "blue"),
    labels = c("Never Treated", "Pre-Treatment")
  )+ 
  labs(
    color = "Treatment Status",
    title = "Public Institution Endowment Pre-Trends", 
    x = "Year",
    y = "Log Mean Endowment"
    )+ theme(text = element_text(size = 20))

ggplot(data = mdf_pre, aes(x = yr, y = f2Endowment, color = as.factor(leg_ever_explicit))) +
  geom_line()+
    scale_color_manual(
    values = c("red", "blue"),
    labels = c("Never Treated", "Pre-Treatment")
)+ 
  labs(
    color = "Treatment Status",
    title = "Private Non-Profit Institution Endowment Pre-Trends", 
    x = "Year",
    y = "Log Mean Endowment"
    ) + theme(text = element_text(size = 20))


# Run actual regression 

m1 <- feols(log(f1AEndowment) ~explicit |  yr , data = mdf_th , cluster = "stateCode")
m2 <- feols(log(f2Endowment) ~explicit |  yr, data = mdf_th , cluster = "stateCode")

m1_c <- feols(log(f1AEndowment) ~explicit  + real_GDP_2017usd + real_personal_income_2017usd + real_PCE_2017usd  |  yr , data = mdf_th , cluster = "stateCode")
m2_c <- feols(log(f2Endowment) ~explicit + real_GDP_2017usd + real_personal_income_2017usd + real_PCE_2017usd  |  yr, data = mdf_th , cluster = "stateCode")

modelsummary(list("Public" = m1,
                  "Private" = m2,
                  "Public (Controls)" = m1_c,
                  "Private (Controls)" = m2_c),
             output = "latex",
             stars = TRUE)


lm_f1a_c <- feols(log_f1AEndowment ~ after_leg_year_explicit + real_GDP_2017usd + real_personal_income_2017usd + real_PCE_2017usd | yr + unitid, data = mdf_th, vcov = "hetero")
lm_f2_c <- feols(log_f2Endowment ~ after_leg_year_explicit + real_GDP_2017usd + real_personal_income_2017usd + real_PCE_2017usd | yr + unitid, data = mdf_th, vcov = "hetero")

lm_f1a <- feols(log_f1AEndowment ~ after_leg_year_explicit| yr + unitid, data = mdf_th, vcov = "hetero")
lm_f2 <- feols(log_f2Endowment ~ after_leg_year_explicit| yr + unitid, data = mdf_th, vcov = "hetero")

modelsummary(list("Public" = lm_f1a, "Private Non-Profit" =lm_f2, "Public (Controls)" = lm_f1a_c, "Private Non-Profit (Controls)" = lm_f2_c), stars=TRUE, output="latex")

# Try dropping always treated 

mdf_drop <- mdf_th %>% 
  filter(!(leg_year < 2004 & explicit == 1))


plotsout_drop <- lapply(log_dep_vars, function(dep){
  ggdid(aggte(att_gt(yname = dep,
                     tname = "yr",
                     idname = "unitid",
                     gname = "leg_year_explicit",
                     xformla = ~1,
                     est_method = "reg",
                     control_group = "notyettreated",
                     data = mdf_drop), type = "dynamic", na.rm=TRUE), title = dep)

})

plotsout_drop[[1]] +
  labs(title = "Group-Time Difference-in-Differences Estimates for Public Institutions", 
        y = "Group Time ATT Estimate", 
        x = "Year Relative to Treatment") +
  theme(
    plot.title = element_text(
      hjust = 0.5,      # centers the title
      size = 18,        # makes it larger
      face = "bold",     # optional: makes it bold
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black")                  # tick labels
  )


plotsout_drop[[2]] +
  labs(title = "Group-Time Difference-in-Differences Estimates for Private Non-Profit Institutions", 
        y = "Group Time ATT Estimate", 
        x = "Year Relative to Treatment") +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      size = 18,
      face = "bold",
      color = "black"
    ),
    axis.title = element_text(color = "black", face = "bold"),  # axis titles
    axis.text  = element_text(color = "black") )                 # tick labels


lm_f1a_cd <- feols(log_f1AEndowment ~ after_leg_year_explicit + real_GDP_2017usd + real_personal_income_2017usd + real_PCE_2017usd | yr + unitid, data = mdf_drop, vcov = "hetero")
lm_f2_cd <- feols(log_f2Endowment ~ after_leg_year_explicit + real_GDP_2017usd + real_personal_income_2017usd + real_PCE_2017usd | yr + unitid, data = mdf_drop, vcov = "hetero")

lm_f1ad <- feols(log_f1AEndowment ~ after_leg_year_explicit| yr + unitid, data = mdf_drop, vcov = "hetero")
lm_f2d <- feols(log_f2Endowment ~ after_leg_year_explicit| yr + unitid, data = mdf_drop, vcov = "hetero")

modelsummary(list("Public" = lm_f1ad, "Private Non-Profit" =lm_f2d, "Public (Controls)" = lm_f1a_cd, "Private Non-Profit (Controls)" = lm_f2_cd), stars=TRUE, output="latex")

# Map state legislation

us_map <- map_data("state")

thesis_treatment %<>% 
  mutate( 
    State = tolower(State)
  )

us_map %<>% 
  left_join(thesis_treatment, by = c("region" = "State"))

us_map %<>% 
  mutate(
    explicit = ifelse(Transparency == "Explicit", TRUE,   FALSE)
  )

ggplot(us_map, aes(long, lat, group = group, fill = Transparency)) +
  geom_polygon(color = "white") +
  coord_fixed(1.3) +
  theme_void() +
  scale_fill_brewer(palette = "Set2") +
  labs(fill = "Category", title = "US States by Treatment Category")  # STATES BY CATEGORY

ggplot(us_map, aes(long, lat, group = group, fill = explicit)) +
  geom_polygon(color = "white") +
  coord_fixed(1.3) +
  theme_void() +
  scale_fill_brewer(palette = "Set2") +
  scale_fill_manual(
    values = c("gray30", "#6666ff"),
    labels = c("Not Treated", "Treated")
  )+ 
  ggtitle("US States by Explicit Treatment Status") +
  theme(plot.title = element_text(hjust = 0.5, size = 60)) +
  labs(
    fill = "Treatment Status"
    ) + theme(legend.text = element_text(size = 20), plot.title = element_text(size = 35)) # STATES BY CATEGORY




# Summary Statistics for Thesis


mdf_th %<>% 
  arrange(unitid, yr) %>% 
  group_by(unitid) %>% 
  mutate(
    endowment_change = endowment - lag(endowment)
  ) %>% 
  ungroup()

summary_table <- mdf_th %>%
  summarise(
    n = n(),
    across(
      .cols = c(endowment, totalAssets, totalRevenue, tuitionAndFees, ugTot, gsTot, admissionsTot, pctAdmitted),
      .fns = list(
        mean = ~mean(.x, na.rm = TRUE),
        sd   = ~sd(.x, na.rm = TRUE),
        min  = ~min(.x, na.rm = TRUE),
        max  = ~max(.x, na.rm = TRUE)
      )
    )
  )

  summary_table

datasummary(
  endowment/1e6 + totalAssets/1e6 + totalRevenue/1e6 + tuitionAndFees/1e6 + ugTot + gsTot + admissionsTot + pctAdmitted + endowment_change/1e6 + leg_ever_explicit +leg_year ~ 
    Mean + SD + Min + Max,
  data = mdf_th,
  output = "latex",
)


test <- mdf_th %>%  
  filter(endowment < 1)

# Plot endowments over time

mdf_plot <- mdf_th %>% 
  group_by(yr) %>% 
  summarise(
    mean_endowment = mean(endowment, na.rm = TRUE)
  )

ggplot(mdf_plot, aes(x = yr, y = mean_endowment/1e6)) +
  geom_line() +
  labs(
    title = "Mean Endowment Over Time",
    x = "Year",
    y = "Mean Endowment (Millions USD)"
  ) +
    theme_minimal()+
  theme(text = element_text(size = 40))



```